---
title: "Sport Harvest Estimator"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    source_code: embed
    theme:
      version: 4
      bootswatch: default
      bg: "#D0D0D0"
      fg: "#000000"
      navbar-bg: "#377F48"
      primary: "#377F48"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", warning = F, echo = F)

library("flexdashboard")
library("tidyverse")
library("shiny")
library("patchwork")
library("gt")

theme_set(theme_light())

pal_wkend <- c(wday = grey(0.3), wend = grey(0.8))

pal_m <- set_names(
  c("#241E33", "#18305A", "#004A70", "#006072", "#007373",
    "#158674", "#4B9976", "#74A97D", "#9CB98A", "#BECA92",
    "#E9DB87", "#FFE56C"), month.abb)

pal_ts <- set_names(
   c("#D2A554", "#626B5D", "#8C8F9E", "#858753", "#A4BADF"),
   paste(1:5, c("Jan-Jun","Jul","Aug","Sep","Oct-Dec"), sep = ":")
)

levels = list(
  dow = c("Thu","Wed","Tue","Mon","Sun","Sat","Fri")
)

lu <- tibble(
  pssp_areas = c("05","06","07","09","10","11","12","13","81","82"),
  fram_ps_spt = c(91:93, 107, 118, 129, 152, 136, 106, 115)
)

out <- reactiveValues(
  est = expand_grid(
    area_code = lu$pssp_areas,
    timestep = as.character(1:5),
    est = 0, 
    wday_df = 0, wend_df = 0,
    wday_apd = 0, wend_apd = 0,
    wday_hapd = 0, wend_hapd = 0,
    rationale = ""
  )
)

load("sport_harvest_estimator.RData")

year_min <- 2009

pssp <- filter(pssp, year >= year_min)
rmis_cs <- filter(rmis_cs, year >= year_min)
```

```{r pssp_data_prep, eval=FALSE}
# #standalone sqlite database from TG supporting estimate report
# #contains enhanced version of pssp_dev mvw_dockside_salmon
# #for coho only but with 2016
# #also has regs datasets with MSF back to 2003; NS currently only back to 2015, but AHB updated externally

db_con <- DBI::dbConnect(RSQLite::SQLite(), 'O:/new_post_season/coho.db')

coho_dockside <- tbl(db_con, "coho_dockside") |>
  filter(
    catch_area_code %in% c("05","06","07","81","82","09","10","11","12","13")
  ) |>
  group_by(area_code = catch_area_code, survey_datetime) |>
  summarise(across(c(anglers, starts_with("coho_")), ~sum(., na.rm=T)), .groups = "drop") |>
  collect() |>
  mutate(
    across(anglers:coho_unk_r, ~as.integer(replace_na(.,0))),
    survey_datetime = as.Date(survey_datetime),
    year = lubridate::year(survey_datetime),
    yr = factor(year),
    month = lubridate::month(survey_datetime),
    m = factor(month.abb[month], levels = month.abb),
    day = lubridate::day(survey_datetime),
    dow = lubridate::wday(survey_datetime, label = T) |> factor(levels = levels$dow),
    wkend = if_else(dow %in% c("Fri", "Sat", "Sun"), "wkend", "wkday"),
    ts = case_when(
      m %in% month.abb[1:6] ~ 1,
      m == "Jul" ~ 2,
      m == "Aug" ~ 3,
      m == "Sep" ~ 4,
      m %in% month.abb[10:12] ~ 5),
    coho_K = coho_ad_k + coho_um_k + coho_unk_k,
    k_ad_rate = coho_ad_k/(coho_ad_k + coho_um_k), #observation specific mark rate of kept
    k_ad_rate = if_else(is.nan(k_ad_rate), 0, k_ad_rate),
    coho_K_ad = coho_ad_k + (coho_unk_k * k_ad_rate),
    coho_R = coho_ad_r + coho_um_r + coho_unk_r,
    coho_E = coho_K + coho_R,
    coho_kpa = coho_K / anglers,
    coho_kpa_ad = coho_K_ad / anglers,
    coho_rpa = coho_R / anglers,
    coho_epa = coho_E / anglers
    ) |>
  select(area_code, survey_datetime, year:ts,
         anglers, coho_K:coho_epa)

DBI::dbDisconnect(db_con); rm(db_con)

#AHB updated from same estimate db, adding NS back to 2003
coho_regulations <- readxl::read_excel(
  path = "T:/DFW-Salmon Mgmt Modeling Team - General/Preseason/Coho/2022/coho_regulations.xlsx",
  sheet = "coho_regulations") |>
  select(catch_area_code, regulation_type_code, start_datetime, end_datetime) |>
  mutate(
    catch_area_code = str_pad(catch_area_code, width = 2, pad = "0"),
    across(contains("datetime"), as.Date)
    )

pssp <- coho_regulations |>
  rowwise() |>
  mutate(
    dockside = list(
      coho_dockside |>
        filter(
          area_code == catch_area_code,
          between(survey_datetime, start_datetime, end_datetime)
        )
    )
  ) |>
  unnest(dockside) |>
  select(-catch_area_code, -start_datetime, -end_datetime)

```

```{r rmis_data_prep, eval=FALSE}
#Sample rates based on RMIS catch-sample query by AHB
rmis_cs <- readxl::read_excel("T:/DFW-Salmon Mgmt Modeling Team - General/Preseason/Coho/2022/SportSampleRateCoho2000_2020.xlsx", sheet = "Summary") |> 
  mutate(
    area_code = str_pad(str_squish(str_remove_all(catch_location_name, "MARINE SPORT AREA |MARINE SPORT PCA|\\.")), width = 2, pad = "0"),
    yr = factor(catch_year)
  ) |> #count(area_code)
  select(FisheryID = `Fish ID`, area_code, year= catch_year, yr, ts = TS, caught = `Sum of number_caught`, sampled = `Sum of number_sampled`, sr = Rate) |> 
  filter(
    !is.na(area_code),
    caught > 10, sampled > 10
    )

```

```{r fram_data_prep, eval=FALSE}
mdb_pre <- "O:/code/coho/fram_mdbs/US_PFMC_NOF_FinalCohoFRAM_MultiYr.mdb"
mdb_pst <- "O:/code/coho/fram_mdbs/PSC_CoTC_PostSeason_CohoFRAMDB_thru2020_021622.mdb"

fs_ps_spt <- bind_rows(
  framr::read_coho_fish_sclr(mdb_pre, fisheries = lu$fram_ps_spt) |> mutate(type = "pre"),
  framr::read_coho_fish_sclr(mdb_pst, fisheries = lu$fram_ps_spt) |> mutate(type = "pst")
  ) |>
  filter(RunYear >= 2009, FisheryFlag != 0) |> #count(type)
  select(type, yr = RunYear, FisheryName, ts = TimeStep, ff = FisheryFlag, Quota, MSFQuota) |>
  mutate(
    area_code = str_remove_all(FisheryName, "Area|Spt|Ar|-") |> str_trim() |> str_pad(width = 2, pad = "0"),
    yr = factor(yr)
  ) |>
  unite("yr_type", yr, type, remove = FALSE) |>
  pivot_longer(names_to = "var", values_to = "val", Quota:MSFQuota) |>
  filter(
    !(ff %in% 1:2 & var=="MSFQuota"),
    !(ff %in% 7:8 & var=="Quota")
    )

```

```{r crc_data_prep, eval=FALSE}
crc <- readRDS("T:/DFW-Salmon Mgmt Modeling Team - General/Catch datasets/CRC/d_spt_hvst_est_20220126.rds") |>
  filter(
    species == "Coho", between(catchyear, 2009, 2022),
    areacode %in% c("5","6","7","81","82","9","10","11","12","13")
    ) |>
  mutate(area_code = str_pad(areacode, width = 2, pad = "0")) |>
  select(area_code, catchyear, catchstatmonth, catchperiodstartdate, catchperiodenddate, catchest) |>
  group_by(area_code, catchyear, catchstatmonth) |> 
  summarise(catchest = sum(catchest), .groups = "drop") |> 
  mutate(
    m = factor(month.abb[catchstatmonth], levels = month.abb),
    yr = factor(catchyear),
    ts = case_when(
      m %in% month.abb[1:6] ~ 1,
      m == "Jul" ~ 2,
      m == "Aug" ~ 3,
      m == "Sep" ~ 4,
      m %in% month.abb[10:12] ~ 5)
    )
```

About
============================================

This application supports the development of preseason coho FRAM marine sport fishery inputs.

The estimated harvest input is defined as the product of days fished $df$, anglers per day $apd$, and harvest-per-angler-per-day $hapd$ or: $harvest_{est} = df * apd * hapd$

Puget Sound dockside creel observations are used to inform the values for anglers per day $apd$, and harvest-per-angler-per-day $hapd$.

Annual sample distribution data are presented when the user selects a focal WDFW Puget Sound catch area and coho FRAM time step from the upper left radio buttons.

The anticipated season length determines the upper $df$ days fished slider, optionally differentiating between weekday and weekend "day types".

The median (q50) and upper range (q95) of the anglers-per-day are used to shape the middle $apd$ slider. The raw values and a sample-rate based expansion are shown, with the slider choices based on latter.

The median and interquartile range (IQR) are used to shape the lower harvest-per-angler-per-day $hapd$ slider. Adjusting the slider modifies the vertical reference line shown in the "PSSP kept obs" plots.

The harvest estimate (i.e., potential FRAM input) resulting from the current slider values is shown in the upper middle green box. 

After reaching an acceptable input, the user enters a brief description of the rationale and clicks the "Keep estimates" button, thereby preserving this text, the input, and the associated underlying $df$, $apd$, and $hapd$. After all desired inputs have been generated, a csv recording each area-timestep can be downloaded.

Coho kept
============================================

Area-Timestep-Estimates {data-height=200}
--------------------------------------------

### Area & Timestep

```{r buttons}
radioButtons("focal_area", label = "Focal Area",
             choices = lu$pssp_areas, selected = lu$pssp_areas[1], inline = T, width = '100%')

radioButtons("focal_ts", label = "Focal Timestep",
             choiceValues = 1:5, 
             choiceNames = paste(1:5, c("Jan-Jun","Jul","Aug","Sep","Oct-Dec"), sep = ":"), 
             selected = 4, inline = T, width = '100%')

```

### Anglers and harvest estimated

```{r valuebox}
renderValueBox({
  valueBox(
    value = paste0("Weekday: ", wday_anglers(), " anglers, ", wday_harvest(), " harvest"),
    icon = "fa-fish")
  })

renderValueBox({
  valueBox(
    value = paste0("Weekend: ", wend_anglers(), " anglers, ", wend_harvest(), " harvest"),
    icon = "fa-fish")
  })

renderValueBox({
  valueBox(
    value = paste0("Total: ",
                   wday_anglers()+wend_anglers(), " anglers, ",
                   wday_harvest()+wend_harvest(), " harvest"),
    icon = "fa-fish")
  })

```

### Output

```{r keep_dl_buttons}
textInput("rationale", label = "Rationale")

actionButton("keep", label = "Keep estimates")

downloadButton("download", label = "Download csv")

```

```{r keep_dl}
reactive({
  est_row <- data.frame(
    area_code = input$focal_area,
    timestep = input$focal_ts,
    est = wday_harvest()+wend_harvest(),
    wday_df = input$wday_df, wend_df = input$wend_df,
    wday_apd = input$wday_apd, wend_apd = input$wend_apd,
    wday_hapd = input$wday_hapd, wend_hapd = input$wend_hapd,
    rationale = input$rationale
    )

  out$est <- rows_update(out$est, est_row, by = c("area_code", "timestep"))
  }) |> 
  bindEvent(input$keep)

downloadHandler(
  filename = function() {
    paste("coho_sport_harvest_est_", Sys.Date(), ".csv", sep="")
  },
  content = function(file) {
    write_csv(out$est, file)
  }
)

```


Controls {data-height=350}
--------------------------------------------

### Weekday assumptions

```{r wkday_sliders}
sliderInput("wday_df", label = "Weekday Days fished",
            min = 0, max = 365, value = 20, step = 1, round = T, width = '100%')

sliderInput("wday_apd", label = "Weekday Anglers per day",
            min = 0, max = 2000, value = 0, step = 5, round = T, width = '100%')

sliderInput("wday_hapd", label = "Weekday Harvest per angler per day",
            min = 0, max = 4, value = 0, step = 0.01, round = F, width = '100%')

wday_harvest <- reactive({
  input$wday_df * input$wday_apd * input$wday_hapd
})

wday_anglers <- reactive({
  input$wday_df * input$wday_apd
})

```

### Weekend assumptions

```{r wkend_sliders}
sliderInput("wend_df", label = "Weekend Days fished",
            min = 0, max = 365, value = 10, step = 1, round = T, width = '100%')

sliderInput("wend_apd", label = "Weekend Anglers per day",
            min = 0, max = 2000, value = 0, step = 5, round = T, width = '100%')

sliderInput("wend_hapd", label = "Weekend Harvest per angler per day",
            min = 0, max = 4, value = 0, step = 0.01, round = F, width = '100%')

wend_harvest <- reactive({
  input$wend_df * input$wend_apd * input$wend_hapd
})

wend_anglers <- reactive({
  input$wend_df * input$wend_apd
})

```


Years {data-height=100}
--------------------------------------------

### Year selection

```{r years_checkboxes}
checkboxGroupInput("years", "Focal years", inline = T, width = '100%',
                   choices = sort(unique(pssp$year)), 
                   selected = sort(unique(pssp$year))
                   )
```

```{r filtered_objects}
pssp_at <- reactive({
  filter(pssp, area_code == input$focal_area, ts == input$focal_ts, yr %in% input$years)
})

rmis_cs_at <- reactive({
  filter(rmis_cs, area_code == input$focal_area, ts == input$focal_ts, yr %in% input$years)
}) 

crc_at <- reactive({
  filter(crc, area_code == input$focal_area, ts == input$focal_ts)
}) 

fram_at <- reactive({
  filter(fs_ps_spt, area_code == input$focal_area, ts == input$focal_ts)
}) 

```

Data {.tabset data-height=800}
--------------------------------------------

### PSSP effort obs (data from *mvw_dockside_salmon*, refreshed Feb 23, 2022)

```{r gt_anglers}
gt::render_gt({
  sr_med <- if_else(nrow(rmis_cs_at()) > 0, median(rmis_cs_at()$sr), NA_real_)
  
  pssp_at_gt <- pssp_at() |> 
    group_by(area_code, yr, wkend) |>
    summarise(
      n_obs = n(),
      across(
        c(anglers),
        list(
          q50 = ~round(quantile(., 0.5, na.rm=T), 0),
          q95 = ~round(quantile(., 0.95, na.rm=T), 0)
        ),
        .names = "{.fn}"), .groups = "drop") |>
    pivot_wider(names_from = wkend, values_from = c(n_obs, q50, q95)) |> 
    arrange(desc(yr)) |>
    left_join(
      rmis_cs_at() |> select(area_code, yr, sr),
      by = c("area_code", "yr")
    ) |> 
    mutate(
      sr = if_else(is.na(sr), sr_med, sr),
      across(
        q50_wkday:q95_wkend,
        list(expnd = ~round(replace_na(.,0) / sr, 0))
      )
    ) |> 
    select(area_code, yr, contains("wkday"), contains("wkend"))
  
  gt::gt(data = pssp_at_gt) |> 
    gt::cols_hide(area_code) |> 
    gt::tab_spanner("Weekday sampled", columns = ends_with("wkday")) |> 
    gt::tab_spanner("Weekday expanded", columns = ends_with("wkday_expnd")) |> 
    gt::tab_spanner("Weekend sampled", columns = ends_with("wkend")) |> 
    gt::tab_spanner("Weekend expanded", columns = ends_with("wkend_expnd")) |> 
    gt::tab_header(
      title = paste("Sampled and expanded anglers/day in Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
      ) |> 
    gt::tab_style(
      style = cell_fill("#D3D3D3"),
      locations = cells_body(columns = c("q50_wkday_expnd", "q50_wkend_expnd"))
    ) |> 
    gt::tab_style(
      style = cell_fill("#FFC46B"),
      locations = cells_body(
        columns = n_obs_wkday:q95_wkday_expnd,
        rows = is.na(n_obs_wkday) | n_obs_wkday <= 5
        )
    ) |> 
    gt::tab_style(
      style = cell_fill("#FFC46B"),
      locations = cells_body(
        columns = n_obs_wkend:q95_wkend_expnd,
        rows = is.na(n_obs_wkend) | n_obs_wkend <= 5
      )
    ) |> 
    gt::summary_rows(
      columns = ends_with("_expnd"),
      fns = list(
        min = ~min(.),
        med = ~median(.),
        max = ~max(.)
      ),
      formatter = fmt_number
    ) |> 
    gt::tab_options(
      table.width = "100%"
    )
  })
```

### PSSP kept obs (data from *mvw_dockside_salmon*, refreshed Feb 23, 2022)

```{r gg_kpa}
renderPlot({
  pssp_at_q <- bind_rows(
    pssp_at() |>  
      pivot_longer(names_to = "kpa_type", values_to = "kpa", cols = c(coho_kpa, coho_kpa_ad)) |> 
      group_by(area_code, yr, wkend, kpa_type) |>
      summarise(
        n_obs = n(),
        across(kpa,
               list(
                 q05 = ~round(quantile(., 0.05, na.rm=T), 2),
                 q25 = ~round(quantile(., 0.25, na.rm=T), 2),
                 q50 = ~round(quantile(., 0.5, na.rm=T), 2),
                 q75 = ~round(quantile(., 0.75, na.rm=T), 2),
                 q95 = ~round(quantile(., 0.95, na.rm=T), 2)
               ),
               .names = "{.fn}"), .groups = "drop") |>
      arrange(desc(yr))
    ,
    pssp_at() |>  
      pivot_longer(names_to = "kpa_type", values_to = "kpa", cols = c(coho_kpa, coho_kpa_ad)) |> 
      group_by(area_code, wkend, kpa_type) |>
      summarise(
        yr = "pooled",
        n_obs = n(),
        across(kpa,
               list(
                 q05 = ~round(quantile(., 0.05, na.rm=T), 2),
                 q25 = ~round(quantile(., 0.25, na.rm=T), 2),
                 q50 = ~round(quantile(., 0.5, na.rm=T), 2),
                 q75 = ~round(quantile(., 0.75, na.rm=T), 2),
                 q95 = ~round(quantile(., 0.95, na.rm=T), 2)
               ),
               .names = "{.fn}"), .groups = "drop")
  )
  
  pssp_at_q |> 
    ggplot(aes(x = yr, y = q50, color = kpa_type, group = kpa_type)) +
    coord_flip() +
    geom_linerange(aes(ymin = q05, ymax = q95), position = position_dodge(width = 0.4)) +
    geom_pointrange(aes(ymin = q25, ymax = q75), size = 1.1, position = position_dodge(width = 0.4)) +
    geom_text(aes(x = yr, y = 0, label = n_obs), nudge_y = -0.05) +
    geom_text(aes(x = yr, y = q05, label = q05), position = position_dodge(width = 0.9)) +
    geom_text(aes(x = yr, y = q50, label = q50), position = position_dodge(width = 1.2)) +
    geom_text(aes(x = yr, y = q95, label = q95), position = position_dodge(width = 0.9)) +
    geom_hline(
      data = tibble(
        wkend = c("wkday", "wkend"), 
        yint = c(input$wday_hapd, input$wend_hapd)
        ),
      aes(yintercept = yint), size = 1.7, color = pal_wkend) +
    facet_wrap(~wkend, scales = "free_x") +
    scale_x_discrete("") +
    scale_y_continuous("Coho kept per angler per day") +
    scale_color_manual(values = c("coho_kpa" = "#D2A554", "coho_kpa_ad" = "#A4BADF")) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
      ) +
    theme(
      text = element_text(size = 20),
      legend.position = "bottom"
      )
})

```

### FRAM

```{r gg_fram}
renderPlot({
  fram_at() |> 
    ggplot(aes(yr_type, val, fill = var)) + 
    geom_col(position = position_stack(), alpha = 0.7) + 
    geom_col(data = filter(fram_at(), type == "pst"), position = position_stack(), alpha = 1) + 
    geom_hline(yintercept = wday_harvest() + wend_harvest()) +
    scale_y_continuous("FRAM inputs", sec.axis = dup_axis(), labels = scales::comma) +
    scale_fill_manual(values = c("Quota" = "#D2A554", "MSFQuota" = "#A4BADF")) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
    ) +
    theme(
      text = element_text(size = 20)
      )
})

```

### CRC estimates

```{r gg_crc}
renderPlot({
  crc_at() |> 
    ggplot(aes(yr, catchest, fill = m)) + 
    geom_col(position = position_stack()) + 
    geom_hline(yintercept = wday_harvest() + wend_harvest()) +
    scale_y_continuous("CRC harvest estimate", sec.axis = dup_axis(), labels = scales::comma) +
    scale_fill_manual(values = pal_m) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
    ) +
    theme(
      text = element_text(size = 20)
      )
})

```

### Estimates

```{r estimate_table}
DT::renderDT({out$est})

```

