---
title: "Sport Harvest Estimator"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    source_code: embed
    theme:
      version: 4
      bootswatch: default
      #bg: "#D0D0D0"
      #fg: "#000000"
      navbar-bg: "#377F48"
      primary: "#377F48"
runtime: shiny
---

```{r setup_libraries, include=FALSE}
knitr::opts_chunk$set(comment = "#>", warning = F, echo = F)

library("flexdashboard")
library("tidyverse")
library("shiny")
library("patchwork")
library("gt")



```

```{r setup_db, include=FALSE}

# this is checking if the application is in a container
# (on the open internet) if it's not,
# it won't connect to the mongo db for storing user inputs
# pure web develop stuff has no influence on fish data


on_web_server = file.exists('/.dockerenv')

if (on_web_server) {
  # load mongo, connect
  library("mongolite")
  user_ests <- mongo(collection = 'user_ests', url = 'mongodb://mongo')
  
  # if saved dataframe doesn't exist yet populate it
  if (nrow(user_ests$find()) == 0) {
    user_ests$insert(
      expand_grid(
        area_code = c("05", "06", "07", "09", "10", "11", "12", "13", "81", "82"),
        timestep = as.character(1:5),
        years = "",
        reg = "",
        est = 0,
        wday_df = 0,
        wend_df = 0,
        wday_hpd = 0,
        wend_hpd = 0,
        rationale = ""
      )
      
    )
  }
}

```


```{r setup_theme, include=FALSE}
theme_set(theme_light())

pal_wkend <- c(wday = grey(0.3), wend = grey(0.8))

pal_m <- set_names(
  c("#241E33", "#18305A", "#004A70", "#006072", "#007373",
    "#158674", "#4B9976", "#74A97D", "#9CB98A", "#BECA92",
    "#E9DB87", "#FFE56C"), month.abb)

pal_ts <- set_names(
   c("#D2A554", "#626B5D", "#8C8F9E", "#858753", "#A4BADF"),
   paste(1:5, c("Jan-Jun","Jul","Aug","Sep","Oct-Dec"), sep = ":")
)

# this is on someone's personal computer 
# and not on a web server this table needs
# to be stored locally
if (on_web_server == FALSE) {
  out <- reactiveValues(
    est = expand_grid(
      area_code = c("05", "06", "07", "09", "10", "11", "12", "13", "81", "82"),
      timestep = as.character(1:5),
      est = 0,
      years = "",
      reg = "",
      wday_df = 0,
      wend_df = 0,
      wday_hpd = 0,
      wend_hpd = 0,
      rationale = ""
    )
  )
}



```


```{r pull_data}
# if on a web server pull data from a saved environment
# if it's not, pull from the sqlite database
if (on_web_server){
  load(file = "sport_harvest_estimator.RData")
} else {
  # connect to database
  con <- DBI::dbConnect(RSQLite::SQLite(), here::here('data/coho_harvest_estimator.db'))
  # pull necessary tables and view
  
  # puget sound creel information merged to regulation
  pssp <- DBI::dbReadTable(con, 'vw_estimates_by_regulation')
  
  
  # puget sound creel information merged to regulation
  regs <- DBI::dbReadTable(con, 'coho_regulations')
  
  
  # compilation of FRAM parameters from post-season and pre-season databases
  fs_ps_spt <- DBI::dbReadTable(con, 'coho_fram')
  
  # disconnect from database
  DBI::dbDisconnect(con)
}

#rm(con)

# regulations need to be split out with weekday and weekday days counted
# for default inputs for days in upcoming fisheries fisheries

daily_regs <- regs %>%
  rowwise() %>%
  mutate(
    date = list(seq(as.Date(start_datetime), as.Date(end_datetime), by = '1 day'))
  ) %>%
  unnest_longer(date) %>%
  select(-start_datetime, -end_datetime) %>% # now a dataframe of daily dates by reg/area
  mutate(
    day_type = if_else(lubridate::wday(date) %in% c(1,6,7), 'WE', 'WD'), # weekday / weekend
    time_step = case_when(
      lubridate::month(date) %in% c(1,2,3,4,5,6) ~ 1L,
      lubridate::month(date)  == 7 ~ 2L,
      lubridate::month(date)  == 8 ~ 3L,
      lubridate::month(date)  == 9 ~ 4L,
      lubridate::month(date) %in% c(10, 11, 12) ~ 5L
      ),
    year = lubridate::year(date)
  )

# ugh, realized dates!!!!!!!!!!!!!!!!!!!!!!!!!!!! need planned/pamphlet!!!
regs_day_type_count <- daily_regs %>%
  count(catch_area_code, regulation_type_code, time_step, day_type, year) %>%
  pivot_wider(names_from = 'day_type', values_from = n)


# this is selecting the most recent number of WD/WE days 
# by area/timestep - defaults of the user interface

recent_regs <- regs_day_type_count %>%
  group_by(catch_area_code, time_step) %>%
  slice_max(order_by = year, n = 1) %>%
  ungroup()

reg_days <- daily_regs %>%
  count(catch_area_code, year, time_step, day_type) %>%
  pivot_wider(names_from = day_type, values_from =n)


```

About
============================================

This application supports the development of preseason coho FRAM marine sport fishery inputs.

The estimated harvest input is defined as the product of days fished $df$, anglers per day $apd$, and harvest-per-angler-per-day $hapd$ or: $harvest_{est} = df * apd * hapd$

Puget Sound dockside creel observations are used to inform the values for anglers per day $apd$, and harvest-per-angler-per-day $hapd$.

Annual sample distribution data are presented when the user selects a focal WDFW Puget Sound catch area and coho FRAM time step from the upper left radio buttons.

The anticipated season length determines the upper $df$ days fished slider, optionally differentiating between weekday and weekend "day types".

The mean, median (q50), and upper range (q95) of the anglers-per-day are used to inform the middle $apd$ slider. The raw values and a sample-rate based expansion are shown, with the slider choices based on latter.

The mean, median and interquartile range (IQR) are used to shape the lower harvest-per-angler-per-day $hapd$ slider. Adjusting the slider modifies the vertical reference line shown in the "PSSP kept obs" plots.

The harvest estimate (i.e., potential FRAM input) resulting from the current slider values is shown in the upper middle green box. 

After reaching an acceptable input, the user enters a brief description of the rationale and clicks the "Keep estimates" button, thereby preserving this text, the input, and the associated underlying $df$, $apd$, and $hapd$. After all desired inputs have been generated, a csv recording each area-timestep can be downloaded.

Coho kept
============================================

Area-Timestep-Estimates {data-height=300}
--------------------------------------------

### Area & Timestep

```{r buttons}
radioButtons("focal_area", label = "Focal Area",
             choices = c("05", "06", "07", "09", "10", "11", "12", "13", "81", "82"), selected = "05", inline = T, width = '100%')

radioButtons("focal_ts", label = "Focal Timestep",
             choiceValues = 1:5, 
             choiceNames = paste(1:5, c("Jan-Jun","Jul","Aug","Sep","Oct-Dec"), sep = ":"), 
             selected = 4, inline = T, width = '100%')

checkboxGroupInput("msf", label = "Regulation",
             choices = c('NS', 'MSF'), 
             selected = c('NS','MSF'), inline = T, width = '100%')



```

### Harvest estimates

```{r valuebox}
renderValueBox({
  valueBox(
    value = paste0("Weekday: ", round(wday_harvest()), " harvest"),
    icon = "fa-fish")
  })

renderValueBox({
  valueBox(
    value = paste0("Weekend: ", round(wend_harvest()), " harvest"),
    icon = "fa-fish")
  })

renderValueBox({
  valueBox(
    value = paste0("Total: ",
                   round(wday_harvest()+wend_harvest()), " harvest"),
    icon = "fa-fish")
  })

```

### Output

```{r keep_dl_buttons}


textInput("rationale", label = "Rationale")
actionButton("keep", label = "Keep estimates")

```

```{r keeper, echo=F}
observeEvent(
  input$keep,{
  est_row <- data.frame(
     area_code = input$focal_area,
     timestep = input$focal_ts,
     est = wday_harvest()+wend_harvest(),
     years = paste0(input$years, collapse = ','),
     reg = paste0(input$msf, collapse = ','),
     wday_df = input$wday_df, wend_df = input$wend_df,
     wday_hpd = input$wday_hpd, wend_hpd = input$wend_hpd,
     rationale = input$rationale
     )
  
  if (on_web_server){
    # pull df out of mongo
    est <- user_ests$find()
    
    # insert modified df
    est_insert <- rows_update(est, est_row, by = c("area_code", "timestep"))
    
    #remove previous data
    user_ests$remove('{}')
    
    #put new data in 
    user_ests$insert(est_insert)
    
  } else {
    # if not a webserver use local
    out$est <- rows_update(out$est, est_row, by = c("area_code", "timestep"))
  }

  updateTextInput(inputId = 'rationale', value = '')
  showNotification('Input saved', type = 'message')
   
})


```


```{r filtered_objects}
# these block of code will make dataframes reactive objects.
# i.e. when the variable pssp_at() is called farther down
# this code it's referencing the reactive object below,
# as the user changes the inputs of area, timestep, and years
# the pssp_at() object executes the code below, dynamically filtering the
# pssp dataframe to the users input represented by pssp_at() 

pssp_at <- reactive({
  # no base period area 12 for july, add july to august
  filter(pssp, 
         area_code == input$focal_area,
         time_step == input$focal_ts, 
         year %in% input$years)
})

fram_at <- reactive({
    filter(fs_ps_spt,
         area_code == input$focal_area, 
         ts == input$focal_ts, 
         yr %in% input$years
    )
}) 

recent_regs_at <- reactive({
  filter(recent_regs,
         catch_area_code == input$focal_area, 
         time_step == input$focal_ts)
})

regs_at <- reactive({
  filter(reg_days, catch_area_code == input$focal_area, 
         time_step == input$focal_ts)
})

# next year move some of these filters to the individual charts / whatever code
# likely losing performance as-is
pssp_all_years <- reactive({
  filter(pssp, 
         area_code == input$focal_area,
         time_step == input$focal_ts
  )
})

fram_all_years <- reactive({
    filter(fs_ps_spt,
         area_code == input$focal_area, 
         ts == input$focal_ts
    )
}) 

```



```{r default_controls}
# these are the controls for the defaults for the weekday and weekend
# inputs. the days are pulled from the most recent fishery by area/timestep
# harvest is the mean selected from the year checkboxes


# application supplied harvest per day
observeEvent(ignoreInit = TRUE,
             c(input$focal_area,
               input$focal_ts,
               input$years,
               input$msf),
             {
               # catch a corner case, if user input doesn't have fisheries to
               # take a mean of, turn off the regulation filtering
               if (nrow(pssp_at()  %>% filter(regulation_type_code %in% input$msf)) == 0) {
                 showModal(
                   modalDialog(
                     title = 'No Usuable Data',
                     'There are no previous data to calculate mean harvest by day.
      Defaulting to whatever data available'
                   )
                 )
                 
                 updateCheckboxGroupInput(inputId = 'msf', selected = c('NS', 'MSF'))
                 updateCheckboxGroupInput(inputId = 'years', selected = 2003:2023)
               } else {
                 if (on_web_server) {
                   saved_input <- user_ests$find(
                     glue::glue(
                       '{"area_code": "{{input$focal_area}}", "timestep": "{{input$focal_ts}}"}',
                       .open = "{{",
                       .close = "}}"
                     )
                   )
                   
                   saved_wday <- saved_input %>%
                     .$wday_df
                   
                   saved_wend <- saved_input %>%
                     .$wend_df
                   
                   saved_wday_h <- saved_input %>%
                     .$wday_hpd
                   
                   saved_wend_h <- saved_input %>%
                     .$wend_hpd
                   
                   saved_years <- saved_input %>%
                     .$years
                   
                   saved_years <- strsplit(saved_years, ",")[[1]]
                   
                   saved_regs <- saved_input %>%
                     .$reg
                   
                   saved_regs <- strsplit(saved_regs, ",")[[1]]
                   
                   saved_ration <- saved_input %>%
                      .$rationale
                   
                   if (length(saved_years) == 0 | !setequal(saved_years, input$years) | !setequal(saved_regs, input$msf)) {
                     mean_harvest <- pssp_at() |>
                       filter(regulation_type_code %in% input$msf) |>
                       rowwise() %>%
                       mutate(total_retained_coho = across(starts_with('coho'), sum)) |>
                       group_by(year, time_step, area_code, day_type) %>%
                       summarize(
                         total_retained_coho = sum(total_retained_coho),
                         days = n(),
                         .groups = 'drop'
                       ) %>%
                       mutate(coho_per_day = total_retained_coho / days) %>%
                       group_by(time_step, area_code, day_type) %>%
                       summarize(mean_daily_harvest = mean(coho_per_day, na.rm = T),
                                 .groups = 'drop') %>%
                       pivot_wider(names_from = day_type, values_from  = mean_daily_harvest) %>%
                       ungroup()
                     
                     
                     
                     
                     wd_mean_harvest <- mean_harvest %>%
                       .$WD
                     
                     we_mean_harvest <- mean_harvest %>%
                       .$WE
                     
                     updateNumericInput(inputId = 'wend_hpd',
                                        value = round(we_mean_harvest))
                     
                     updateNumericInput(inputId = 'wday_hpd',
                                        value = round(wd_mean_harvest))
                     
                   } else{
                     
                     updateNumericInput(inputId = 'wday_df', value = saved_wday)
                     updateNumericInput(inputId = 'wend_df', value = saved_wend)
                     
                     
                     updateNumericInput(inputId = 'wend_hpd', value = saved_wend_h)
                     updateNumericInput(inputId = 'wday_hpd', value = saved_wday_h)
                     
                     
                     updateCheckboxGroupInput(inputId = 'years', selected = as.numeric(saved_years))
                     updateCheckboxGroupInput(inputId = 'msf', selected = saved_regs)
                     
                     updateTextInput(inputId = 'rationale', value =  saved_ration)
                     
                   }
                   
                 } else {
                   mean_harvest <- pssp_at() |>
                     filter(regulation_type_code %in% input$msf) |>
                     rowwise() %>%
                     mutate(total_retained_coho = across(starts_with('coho'), sum)) |>
                     group_by(year, time_step, area_code, day_type) %>%
                     summarize(
                       total_retained_coho = sum(total_retained_coho),
                       days = n(),
                       .groups = 'drop'
                     ) %>%
                     mutate(coho_per_day = total_retained_coho / days) %>%
                     group_by(time_step, area_code, day_type) %>%
                     summarize(mean_daily_harvest = mean(coho_per_day, na.rm = T),
                               .groups = 'drop') %>%
                     pivot_wider(names_from = day_type, values_from  = mean_daily_harvest) %>%
                     ungroup()
                   
                   
                   
                   
                   wd_mean_harvest <- mean_harvest %>%
                     .$WD
                   
                   we_mean_harvest <- mean_harvest %>%
                     .$WE
                   
                   updateNumericInput(inputId = 'wend_hpd', value = round(we_mean_harvest))
                   
                   updateNumericInput(inputId = 'wday_hpd', value = round(wd_mean_harvest))
                   
                   
                 }
                 
               }
               
             })



# application supplied weekend/weekdays
observeEvent(ignoreInit = TRUE,
             c(input$focal_area, input$focal_ts), {

               
               # # pull from database if there is a data saved
               # if (on_web_server) {
               #   saved_input <- user_ests$find(
               #     glue::glue(
               #       '{"area_code": "{{input$focal_area}}", "timestep": "{{input$focal_ts}}"}',
               #       .open = "{{",
               #       .close = "}}"
               #     )
               #   )
               #   
               #   
               #   saved_wday <- saved_input %>%
               #     .$wday_df
               #   
               #   saved_wend <- saved_input %>%
               #     .$wend_df
               #   
               #   saved_wday_h <- saved_input %>%
               #     .$wday_hpd
               #   
               #   saved_wend_h <- saved_input %>%
               #     .$wend_hpd
               #   
               #   saved_years <- saved_input %>%
               #     .$years
               #   
               #   saved_years <- strsplit(saved_years, ",")[[1]]
               #   
               #   saved_regs <- saved_input %>%
               #     .$reg
               #   
               #   saved_regs <- strsplit(saved_regs, ",")[[1]]
               #   
               #   saved_ration <- saved_input %>%
               #   .$rationale
               #   
               #   # if nothing as been saved
               #   
               #   if (length(saved_years) == 0) {
               #     wd_days <- recent_regs_at() %>%
               #       # calculate most recent regulation weedays
               #       .$WD
               #     
               #     updateNumericInput(inputId = 'wday_df', value = wd_days)
               #     
               #     
               #     we_days <- recent_regs_at() %>%
               #       .$WE
               #     
               #     updateNumericInput(inputId = 'wend_df', value = we_days)
               #     
               #   } else {
               #     # set days from database
               #     updateNumericInput(inputId = 'wday_df', value = saved_wday)
               #     updateNumericInput(inputId = 'wend_df', value = saved_wend)
               #     
               #     
               #     updateNumericInput(inputId = 'wend_hpd', value = saved_wend_h)
               #     updateNumericInput(inputId = 'wday_hpd', value = saved_wday_h)
               #     
               #     
               #     updateCheckboxGroupInput(inputId = 'years', selected = as.numeric(saved_years))
               #     updateCheckboxGroupInput(inputId = 'msf', selected = saved_regs)
               #     
               #     updateTextInput(inputId = 'rationale', value =  saved_ration)
               #     
               #   }
               #   
               #   
               #   
               # } else {
               #   wd_days <- recent_regs_at() %>%
               #     # calculate most recent regulation weekdays
               #     .$WD
               #   
               #   updateNumericInput(inputId = 'wday_df', value = wd_days)
               #   
               #   
               #   we_days <- recent_regs_at() %>%
               #     .$WE
               #   
               #   updateNumericInput(inputId = 'wend_df', value = we_days)
               # }
               
                 wd_days <- recent_regs_at() %>%
                   # calculate most recent regulation weekdays
                   .$WD
                 
                 updateNumericInput(inputId = 'wday_df', value = wd_days)
                 
                 
                 we_days <- recent_regs_at() %>%
                   .$WE
                 
                 updateNumericInput(inputId = 'wend_df', value = we_days)
               
})





```


Controls {data-height=225}
--------------------------------------------

### Weekday assumptions
```{r wkday_sliders}



numericInput("wday_df", label = "Weekday Days fished",
            min = 0, max = 365, value = 20, step = 1, width = '100%')

# numericInput("wday_apd", label = "Weekday Anglers per day",
#             min = 0, max = 2000, value = 0, step = 5, width = '100%')

numericInput("wday_hpd", label = "Mean Weekday Day Harvest",
            min = 0, max = 4, value = 0, step = 0.01, width = '100%')

wday_harvest <- reactive({
  input$wday_df * input$wday_hpd
})

```

### Weekend assumptions

```{r wkend_sliders}
numericInput("wend_df", label = "Weekend Days fished",
            min = 0, max = 365, value = 20, step = 1, width = '100%')

# numericInput("wend_apd", label = "Weekend Anglers per day",
#             min = 0, max = 2000, value = 0, step = 5,  width = '100%')

numericInput("wend_hpd", label = "Mean Weekend Day Harvest",
            min = 0, max = 4, value = 0, step = 0.01,  width = '100%')

wend_harvest <- reactive({
  input$wend_df * input$wend_hpd
})


```





Years {data-height=125}
--------------------------------------------

### Year selection

```{r years_checkboxes}
actionLink('select_all_years', 'Select All')
actionLink('select_recent_years', 'Select Recent 6 years')

checkboxGroupInput("years", "Focal years", inline = T, width = '100%',
                   choices = 2004:2023,
                   selected = 2004:2023# don't hardcode next year, next year is here and i still hardcoded it bahaha!!!
                   )

observeEvent(
  ignoreInit = TRUE,
  input$select_all_years,
  {
    updateCheckboxGroupInput(inputId = 'years', selected = 2004:2023)
    
    }
)

observeEvent(
  ignoreInit = TRUE,
  input$select_recent_years,
  {
    updateCheckboxGroupInput(inputId = 'years', selected = 2016:2023)
    }
)

```


Data {.tabset data-height=800}
--------------------------------------------

### FRAM
```{r gg_fram}
renderPlot({
  fram_ests <- pssp_at() |>
    filter(regulation_type_code %in% input$msf) |>
    group_by(year, time_step, area_code, source) |>
    summarize(across(starts_with('coho'), sum, na.rm=T), .groups = 'drop') |>
    rowwise() |>
    mutate(
      # apportion unk mark-status to ad/um for retention
      coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / sum(coho_104_ad_ret,coho_104_um_ret , na.rm=T))),
      #coho_ad_k = replace_na(coho_104_ad_ret, 0),
      coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret / sum(coho_104_ad_ret,coho_104_um_ret, na.rm=T))),
      #coho_um_k = replace_na(coho_104_um_ret, 0),
      total_retained_coho = sum(coho_ad_k, coho_um_k),
      yr_type = paste0(year, '_pst_ests'),
      yr = as.character(year),
      ts = time_step,
      type = 'ests'
    ) |>
    select(yr_type, yr, ts, type, area_code, var = source, val = total_retained_coho)     |>
    bind_rows(
      fram_at() %>%
        mutate(
          selector = if_else(var == 'MSFQuota', 'MSF', 'NS') 
        ) %>%
        filter(
          selector %in% input$msf
        ) %>% 
        select(-selector) 
    )
  
  
  
  fram_ests %>%
    ggplot(aes(yr_type, val, fill = var)) +
    geom_col(position = position_stack(), alpha = 0.7) +
    #geom_col(data = filter(fram_ests, type == "ests"), position = position_stack(), alpha = .6) +
    #geom_col(data = filter(fram_ests, type == "pst"), position = position_stack(), alpha = 1) +
    geom_hline(yintercept = wday_harvest() + wend_harvest()) +
    scale_y_continuous("FRAM inputs", sec.axis = dup_axis(), labels = scales::comma) +
    scale_fill_manual(values = c(
      "Quota" = "#D2A554",
      "MSFQuota" = "#A4BADF",
      "CRC" = "#04a43c",
      "CREEL" = "#014218"
    )) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
    ) +
    theme(
      text = element_text(size = 20),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
    )
})

```


### Input vs. Historical
```{r fram_vs_historical}
renderPlot({
  fram_ests <- pssp_all_years() |>
    filter(regulation_type_code %in% input$msf,
           year >= 2010) |>
    group_by(year, time_step, area_code, source) |>
    summarize(across(starts_with('coho'), sum, na.rm=T), .groups = 'drop') |>
    rowwise() |>
    mutate(
      # apportion unk mark-status to ad/um for retention
      coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / sum(coho_104_ad_ret,coho_104_um_ret , na.rm=T))),
      #coho_ad_k = replace_na(coho_104_ad_ret, 0),
      coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret / sum(coho_104_ad_ret,coho_104_um_ret, na.rm=T))),
      #coho_um_k = replace_na(coho_104_um_ret, 0),
      total_retained_coho = sum(coho_ad_k, coho_um_k),
      yr_type = paste0(year, '_pst_ests'),
      yr = as.character(year),
      ts = time_step,
      type = 'ests'
    ) |>
    select(yr_type, yr, ts, type, area_code, var = source, val = total_retained_coho)     |>
    bind_rows(
      fram_all_years() %>%
        mutate(
          selector = if_else(var == 'MSFQuota', 'MSF', 'NS') 
        ) %>%
        filter(
          selector %in% input$msf,
          yr >= 2010
        ) %>% 
        select(-selector) 
    )
  
  
  
  fram_ests %>%
    ggplot(aes(yr_type, val, fill = var)) +
    geom_col(position = position_stack(), alpha = 0.7) +
    #geom_col(data = filter(fram_ests, type == "ests"), position = position_stack(), alpha = .6) +
    #geom_col(data = filter(fram_ests, type == "pst"), position = position_stack(), alpha = 1) +
    geom_hline(yintercept = wday_harvest() + wend_harvest()) +
    scale_y_continuous("FRAM inputs", sec.axis = dup_axis(), labels = scales::comma) +
    scale_fill_manual(values = c(
      "Quota" = "#D2A554",
      "MSFQuota" = "#A4BADF",
      "CRC" = "#04a43c",
      "CREEL" = "#014218"
    )) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
    ) +
    theme(
      text = element_text(size = 20),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
    )
})

```


### FRAM Performance
```{r fram_performance}
renderPlot({
  ests <- pssp_at() %>%
    mutate(
      # take care of NA's
      across(starts_with('coho'), ~replace_na(.x, 0)),
      # apportion unk mark-status to ad/um for retention
      coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / (coho_104_ad_ret + coho_104_um_ret ))),
      coho_ad_k = replace_na(coho_104_ad_ret, 0),
      coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret /  (coho_104_ad_ret + coho_104_um_ret ))),
      coho_um_k = replace_na(coho_104_um_ret, 0),
      total_retained_coho = coho_ad_k + coho_um_k
    ) |>
    group_by(time_step, area_code, year) %>%
    summarize(total_retained_coho = sum(total_retained_coho, na.rm=T), .groups = 'drop')
  
  fram <- fram_at() %>%
    mutate(
      selector = if_else(var == 'MSFQuota', 'MSF', 'NS') 
    ) %>%
    filter(
      selector %in% input$msf
    ) %>% 
    select(-selector) %>%
    mutate(
      year = as.integer(yr),
      time_step = ts
    ) %>%
    group_by(year, time_step, area_code) %>%
    summarize(predicted_retained_coho = sum(val, na.rm=T), .groups = 'drop')
  
  plot <- fram %>%
    inner_join(ests, by = c("year", "time_step", "area_code")) %>%
    mutate(
      percent = (total_retained_coho - predicted_retained_coho) / total_retained_coho,
      ape = abs(percent),
      type = percent < 0,
      sq_dev = (total_retained_coho - predicted_retained_coho)^2,
      
    ) 
  
  rmse <- round(sqrt(plot$sq_dev))
  mape <- round(mean(plot$ape), 2)
  mpe <- round(mean(plot$percent), 2)
  
  plot %>%
    ggplot(aes(factor(year), percent, fill = type)) +
    geom_col(alpha = 0.8) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","), 'RMSE: ', rmse, ' MAPE:', mape, 'MPE: ', mpe, '      (post-pre) / post'),
      x = NULL,
      y = 'Percent Error'
    ) +
    scale_y_continuous(labels = scales::percent) +
    theme(
      axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1),
      text = element_text(size = 20),
      legend.position = 'none'
    )
  
  
})
```

### Daily Harvest
```{r gg_daily}
renderPlot({
  day_plot <- pssp_at() |>
    filter(regulation_type_code %in% input$msf) |>
      mutate(
        # take care of NA's
        across(starts_with('coho'), ~replace_na(.x, 0)),
        # apportion unk mark-status to ad/um for retention
        coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / (coho_104_ad_ret + coho_104_um_ret ))),
        coho_ad_k = replace_na(coho_104_ad_ret, 0),
        coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret /  (coho_104_ad_ret + coho_104_um_ret ))),
        coho_um_k = replace_na(coho_104_um_ret, 0),
        total_retained_coho = coho_ad_k + coho_um_k
      )
  
  day_plot |> 
    ggplot(aes(factor(year), total_retained_coho, fill = day_type)) +
    geom_boxplot() +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ",")),
      y = 'Total Retained Coho',
      x = 'Day Type'
    ) + 
    theme(
      text = element_text(size = 20)
      )
    
})
```

### Mean Daily Harvest

```{r mean_daily_harvest}
# renderText({
#   
# })
# 

renderPlot({
  
  mean_harvest_plot <- pssp_at() |>
    filter(regulation_type_code %in% input$msf) |>
    rowwise() %>%
    mutate(total_retained_coho = coho_104_ad_ret + coho_104_um_ret + coho_104_uk_ret) |>
    group_by(year, regulation_type_code) |>
    summarize(mean_est = mean(total_retained_coho, na.rm = T),
              .groups = 'drop')
  
  input_mean <- ((input$wend_df * input$wend_hpd) +  (input$wday_df * input$wday_hpd)) / (input$wday_df + input$wend_df)
  
  mean_harvest_plot |> 
    ggplot(aes(factor(year), mean_est, fill = regulation_type_code, group=mean_est)) +
    geom_col(position = position_dodge(), alpha=0.8) +
    geom_hline( yintercept = input_mean) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ",")),
      y = 'Mean Coho Harvest Per Day',
      x = 'Year'
    ) +
    theme(
      text = element_text(size = 20)
      )
    
})


#((input$wend_df * input$wend_hpd) +  (input$wday_df * input$wday_hpd)) / (input$wday_df + input$wend_df)




```

### Historical Days Open
```{r regs_table}
renderTable({
  regs_at() %>%
    mutate(total_days = WE + WD) %>%
    arrange(-year) %>%
    mutate(year = as.character(year))
})
```

### Saved Inputs
```{r, download_link}
downloadLink("download", label = "Download csv")
if (on_web_server){
  output$download <- downloadHandler(
    filename = function() {
      paste("coho_sport_harvest_est_", Sys.Date(), ".csv", sep="")
    },
    content = function(file) {
      write_csv(user_ests$find(), file)
    }
  )
} else {
  output$download <- downloadHandler(
    filename = function() {
      paste("coho_sport_harvest_est_", Sys.Date(), ".csv", sep="")
    },
    content = function(file) {
      write_csv(out$est, file)
    }
  )
}

```

```{r ests_table}
if (on_web_server) {
  # poll database
  users_inputs <- reactivePoll(
    1000,
    session,
    
    # check if something changed
    checkFunc = function() {
      ests <- user_ests$find()
      
      
      sum(ests$est,
          ests$wday_df,
          ests$wend_df,
          ests$wday_hpd,
          ests$wend_hpd)
    },
    valueFunc = function() {
      # pull new data
      user_ests$find()
    }
    
    
  )
  # display table from database
  renderTable({
    users_inputs()
  })
} else{
  # otherwise display local
  renderTable({
    out$est
  })
}


```

```{css est_table}
#section-saved-inputs
  .chart-shim {
    overflow-y: scroll;
    }
```

<!-- ### Load Saved Inputs -->

```{r load_inputs}
# still working on this

# renderUI({
#   selectInput("area",label = "Area",
#       choices =as.list(out$est %>%
#                          filter(est > 0) %>%
#                          .$area_code))
# })
# 
# renderUI({
#   selectInput("area",label = "Area",
#       choices =as.list(out$est %>%
#                          filter(est > 0) %>%
#                          .$area_code))
# })


```

