---
title: "Sport Harvest Estimator"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    source_code: embed
    theme:
      version: 4
      bootswatch: default
      bg: "#D0D0D0"
      fg: "#000000"
      navbar-bg: "#377F48"
      primary: "#377F48"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", warning = F, echo = F)

library("flexdashboard")
library("tidyverse")
library("shiny")
library("patchwork")
library("gt")

theme_set(theme_light())

pal_wkend <- c(wday = grey(0.3), wend = grey(0.8))

pal_m <- set_names(
  c("#241E33", "#18305A", "#004A70", "#006072", "#007373",
    "#158674", "#4B9976", "#74A97D", "#9CB98A", "#BECA92",
    "#E9DB87", "#FFE56C"), month.abb)

pal_ts <- set_names(
   c("#D2A554", "#626B5D", "#8C8F9E", "#858753", "#A4BADF"),
   paste(1:5, c("Jan-Jun","Jul","Aug","Sep","Oct-Dec"), sep = ":")
)


out <- reactiveValues(
  est = expand_grid(
    area_code = c("05", "06", "07", "09", "10", "11", "12", "13", "81", "82"),
    timestep = as.character(1:5),
    est = 0, 
    wday_df = 0, wend_df = 0,
    wday_apd = 0, wend_apd = 0,
    wday_hapd = 0, wend_hapd = 0,
    rationale = ""
  )
)
```


```{r pull_data}
# connect to database
con <- DBI::dbConnect(RSQLite::SQLite(), here::here('data/coho_harvest_estimator.db'))
# pull necessary tables and view

# puget sound creel information merged to regulation
pssp <- DBI::dbReadTable(con, 'vw_coho_dockside_regulation')

# sample rates from RMIS catch sample tables
rmis_cs <- DBI::dbReadTable(con, 'coho_sample_rate')

# compilation of FRAM parameters from post-season and pre-season databases
fs_ps_spt <- DBI::dbReadTable(con, 'coho_fram')

# CRC estimates
crc <- DBI::dbReadTable(con, 'coho_crc')

# disconnect from
DBI::dbDisconnect(con)

```

About
============================================

This application supports the development of preseason coho FRAM marine sport fishery inputs.

The estimated harvest input is defined as the product of days fished $df$, anglers per day $apd$, and harvest-per-angler-per-day $hapd$ or: $harvest_{est} = df * apd * hapd$

Puget Sound dockside creel observations are used to inform the values for anglers per day $apd$, and harvest-per-angler-per-day $hapd$.

Annual sample distribution data are presented when the user selects a focal WDFW Puget Sound catch area and coho FRAM time step from the upper left radio buttons.

The anticipated season length determines the upper $df$ days fished slider, optionally differentiating between weekday and weekend "day types".

The mean, median (q50), and upper range (q95) of the anglers-per-day are used to inform the middle $apd$ slider. The raw values and a sample-rate based expansion are shown, with the slider choices based on latter.

The mean, median and interquartile range (IQR) are used to shape the lower harvest-per-angler-per-day $hapd$ slider. Adjusting the slider modifies the vertical reference line shown in the "PSSP kept obs" plots.

The harvest estimate (i.e., potential FRAM input) resulting from the current slider values is shown in the upper middle green box. 

After reaching an acceptable input, the user enters a brief description of the rationale and clicks the "Keep estimates" button, thereby preserving this text, the input, and the associated underlying $df$, $apd$, and $hapd$. After all desired inputs have been generated, a csv recording each area-timestep can be downloaded.

Coho kept
============================================

Area-Timestep-Estimates {data-height=200}
--------------------------------------------

### Area & Timestep

```{r buttons}
radioButtons("focal_area", label = "Focal Area",
             choices = c("05", "06", "07", "09", "10", "11", "12", "13", "81", "82"), selected = "05", inline = T, width = '100%')

radioButtons("focal_ts", label = "Focal Timestep",
             choiceValues = 1:5, 
             choiceNames = paste(1:5, c("Jan-Jun","Jul","Aug","Sep","Oct-Dec"), sep = ":"), 
             selected = 4, inline = T, width = '100%')

```

### Anglers and harvest estimated

```{r valuebox}
renderValueBox({
  valueBox(
    value = paste0("Weekday: ", round(wday_anglers()), " anglers, ", round(wday_harvest()), " harvest"),
    icon = "fa-fish")
  })

renderValueBox({
  valueBox(
    value = paste0("Weekend: ", round(wend_anglers()), " anglers, ", round(wend_harvest()), " harvest"),
    icon = "fa-fish")
  })

renderValueBox({
  valueBox(
    value = paste0("Total: ",
                   round(wday_anglers()+wend_anglers()), " anglers, ",
                   round(wday_harvest()+wend_harvest()), " harvest"),
    icon = "fa-fish")
  })

```

### Output

```{r keep_dl_buttons}
textInput("rationale", label = "Rationale")

actionButton("keep", label = "Keep estimates")

downloadButton("download", label = "Download csv")

```

```{r keep_dl}
reactive({
  est_row <- data.frame(
    area_code = input$focal_area,
    timestep = input$focal_ts,
    est = wday_harvest()+wend_harvest(),
    wday_df = input$wday_df, wend_df = input$wend_df,
    wday_apd = input$wday_apd, wend_apd = input$wend_apd,
    wday_hapd = input$wday_hapd, wend_hapd = input$wend_hapd,
    rationale = input$rationale
    )

  out$est <- rows_update(out$est, est_row, by = c("area_code", "timestep"))
  }) |> 
  bindEvent(input$keep)

downloadHandler(
  filename = function() {
    paste("coho_sport_harvest_est_", Sys.Date(), ".csv", sep="")
  },
  content = function(file) {
    write_csv(out$est, file)
  }
)

```


Controls {data-height=350}
--------------------------------------------

### Weekday assumptions

```{r wkday_sliders}
sliderInput("wday_df", label = "Weekday Days fished",
            min = 0, max = 365, value = 20, step = 1, round = T, width = '100%')

sliderInput("wday_apd", label = "Weekday Anglers per day",
            min = 0, max = 2000, value = 0, step = 5, round = T, width = '100%')

sliderInput("wday_hapd", label = "Weekday Harvest per angler per day",
            min = 0, max = 4, value = 0, step = 0.01, round = F, width = '100%')

wday_harvest <- reactive({
  input$wday_df * input$wday_apd * input$wday_hapd
})

wday_anglers <- reactive({
  input$wday_df * input$wday_apd
})

```

### Weekend assumptions

```{r wkend_sliders}
sliderInput("wend_df", label = "Weekend Days fished",
            min = 0, max = 365, value = 10, step = 1, round = T, width = '100%')

sliderInput("wend_apd", label = "Weekend Anglers per day",
            min = 0, max = 2000, value = 0, step = 5, round = T, width = '100%')

sliderInput("wend_hapd", label = "Weekend Harvest per angler per day",
            min = 0, max = 4, value = 0, step = 0.01, round = F, width = '100%')

wend_harvest <- reactive({
  input$wend_df * input$wend_apd * input$wend_hapd
})

wend_anglers <- reactive({
  input$wend_df * input$wend_apd
})

```


Years {data-height=100}
--------------------------------------------

### Year selection

```{r years_checkboxes}
checkboxGroupInput("years", "Focal years", inline = T, width = '100%',
                   choices = sort(unique(pssp$year)), 
                   selected = sort(unique(pssp$year))
                   )
```

```{r filtered_objects}
pssp_at <- reactive({
  filter(pssp, area_code == input$focal_area, time_step == input$focal_ts, yr %in% input$years)
})

rmis_cs_at <- reactive({
  filter(rmis_cs, area_code == input$focal_area, ts == input$focal_ts, yr %in% input$years)
}) 

crc_at <- reactive({
  filter(crc, area_code == input$focal_area, ts == input$focal_ts)
}) 

fram_at <- reactive({
  filter(fs_ps_spt, area_code == input$focal_area, ts == input$focal_ts)
}) 

```

Data {.tabset data-height=800}
--------------------------------------------

### PSSP effort obs (data from *mvw_dockside_salmon*, refreshed Nov 22, 2022)

```{r gt_anglers}
gt::render_gt({
  sample_rate_median <- if_else(nrow(rmis_cs_at()) > 0, median(rmis_cs_at()$sample_rate), NA_real_)
  
  pssp_at_gt <- pssp_at() |> 
    group_by(area_code, yr, week_day_type) |>
    summarise(
      n_obs = n(),
      across(
        c(anglers),
        list(
          q50 = ~round(quantile(., 0.5, na.rm=T), 0),
          mu = ~round(mean(., na.rm=T), 0),
          q95 = ~round(quantile(., 0.95, na.rm=T), 0)
        ),
        .names = "{.fn}"), .groups = "drop") |>
    pivot_wider(names_from = week_day_type, values_from = c(n_obs, q50, mu, q95)) |> 
    arrange(desc(yr)) |>
    left_join(
      rmis_cs_at() |> select(area_code, yr, sample_rate),
      by = c("area_code", "yr")
    ) |> 
    mutate(
      # this effectively imputes the sample rate
      sample_rate = if_else(is.na(sample_rate), sample_rate_median, sample_rate),
      across(
        q50_wkday:q95_wkend,
        list(expnd = ~round(replace_na(.,0) / sample_rate, 0))
      )
    ) |> 
    select(area_code, yr, contains("wkday"), contains("wkend"))
  
  gt::gt(data = pssp_at_gt) |> 
    gt::cols_hide(area_code) |> 
    gt::tab_spanner("Weekday sampled", columns = ends_with("wkday")) |> 
    gt::tab_spanner("Weekday expanded", columns = ends_with("wkday_expnd")) |> 
    gt::tab_spanner("Weekend sampled", columns = ends_with("wkend")) |> 
    gt::tab_spanner("Weekend expanded", columns = ends_with("wkend_expnd")) |> 
    gt::tab_header(
      title = paste("Sampled and expanded anglers/day in Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
      ) |> 
    gt::tab_style(
      style = cell_fill("#D3D3D3"),
      locations = cells_body(columns = c("q50_wkday_expnd", "q50_wkend_expnd"))
    ) |> 
    gt::tab_style(
      style = cell_fill("#FFC46B"),
      locations = cells_body(
        columns = n_obs_wkday:q95_wkday_expnd,
        rows = is.na(n_obs_wkday) | n_obs_wkday <= 5
        )
    ) |> 
    gt::tab_style(
      style = cell_fill("#FFC46B"),
      locations = cells_body(
        columns = n_obs_wkend:q95_wkend_expnd,
        rows = is.na(n_obs_wkend) | n_obs_wkend <= 5
      )
    ) |> 
    gt::summary_rows(
      columns = ends_with("_expnd"),
      fns = list(
        min = ~min(.),
        med = ~median(.),
        mean = ~mean(.),
        max = ~max(.)
      ),
      formatter = fmt_number
    ) |> 
    gt::tab_options(
      table.width = "100%"
    )
  })
```

### PSSP kept obs (data from *mvw_dockside_salmon*, refreshed Nov 22, 2022)

```{r gg_kpa}
renderPlot({
  pssp_at_q <- bind_rows(
    pssp_at() |>  
      pivot_longer(names_to = "kpa_type",
                   values_to = "kpa", 
                   cols = c(coho_kept_per_angler, coho_kept_ad_per_angler)) |> 
      group_by(area_code, yr, week_day_type, kpa_type) |>
      summarise(
        n_obs = n(),
        across(kpa,
               list(
                 q05 = ~round(quantile(., 0.05, na.rm=T), 2),
                 q25 = ~round(quantile(., 0.25, na.rm=T), 2),
                 q50 = ~round(quantile(., 0.5, na.rm=T), 2),
                 q75 = ~round(quantile(., 0.75, na.rm=T), 2),
                 q95 = ~round(quantile(., 0.95, na.rm=T), 2),
                 mu = ~round(mean(., na.rm=T), 2)
               ),
               .names = "{.fn}"), .groups = "drop") |>
      arrange(desc(yr))
    ,
    pssp_at() |>  
      pivot_longer(names_to = "kpa_type",
                   values_to = "kpa", 
                   cols = c(coho_kept_per_angler, coho_kept_ad_per_angler))|> 
      group_by(area_code, week_day_type, kpa_type) |>
      summarise(
        yr = "pooled",
        n_obs = n(),
        across(kpa,
               list(
                 q05 = ~round(quantile(., 0.05, na.rm=T), 2),
                 q25 = ~round(quantile(., 0.25, na.rm=T), 2),
                 q50 = ~round(quantile(., 0.5, na.rm=T), 2),
                 q75 = ~round(quantile(., 0.75, na.rm=T), 2),
                 q95 = ~round(quantile(., 0.95, na.rm=T), 2),
                 mu = ~round(mean(., na.rm=T), 2)
               ),
               .names = "{.fn}"), .groups = "drop")
  )
  
  pssp_at_q |> 
    ggplot(aes(x = yr, y = q50, color = kpa_type, group = kpa_type)) +
    coord_flip() +
    geom_linerange(aes(ymin = q05, ymax = q95), position = position_dodge(width = 0.4)) +
    geom_pointrange(aes(ymin = q25, ymax = q75), size = 1.1, position = position_dodge(width = 0.4)) +
    geom_point(aes(x = yr, y = mu), shape = 4, size = 6, position = position_dodge(width = 0.4)) +
    geom_text(aes(x = yr, y = 0, label = n_obs), nudge_y = -0.05) +
    geom_text(aes(x = yr, y = q05, label = q05), position = position_dodge(width = 1.1)) +
    geom_text(aes(x = yr, y = q50, label = q50), position = position_dodge(width = 1.3)) +
    geom_text(aes(x = yr, y = mu, label = mu), position = position_dodge(width = 1.5)) +
    geom_text(aes(x = yr, y = q95, label = q95), position = position_dodge(width = 1.1)) +
    geom_hline(
      data = tibble(
        week_day_type = c("wkday", "wkend"), 
        yint = c(input$wday_hapd, input$wend_hapd)
        ),
      aes(yintercept = yint), size = 1.7, color = pal_wkend) +
    facet_wrap(~week_day_type, scales = "free_x") +
    scale_x_discrete("") +
    scale_y_continuous("Coho kept per angler per day") +
    scale_color_manual(values = c("coho_kept_per_angler" = "#D2A554", "coho_kept_ad_per_angler" = "#A4BADF")) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
      ) +
    theme(
      text = element_text(size = 20),
      legend.position = "bottom"
      )
})





```

### FRAM

```{r gg_fram}
renderPlot({
  fram_at() |> 
    ggplot(aes(yr_type, val, fill = var)) + 
    geom_col(position = position_stack(), alpha = 0.7) + 
    geom_col(data = filter(fram_at(), type == "pst"), position = position_stack(), alpha = 1) + 
    geom_hline(yintercept = wday_harvest() + wend_harvest()) +
    scale_y_continuous("FRAM inputs", sec.axis = dup_axis(), labels = scales::comma) +
    scale_fill_manual(values = c("Quota" = "#D2A554", "MSFQuota" = "#A4BADF")) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
    ) +
    theme(
      text = element_text(size = 20)
      )
})

```

### CRC estimates

```{r gg_crc}
renderPlot({
  crc_at() |> 
    ggplot(aes(yr, catchest, fill = m)) + 
    geom_col(position = position_stack()) + 
    geom_hline(yintercept = wday_harvest() + wend_harvest()) +
    scale_y_continuous("CRC harvest estimate", sec.axis = dup_axis(), labels = scales::comma) +
    scale_fill_manual(values = pal_m) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
    ) +
    theme(
      text = element_text(size = 20)
      )
})

```

### Estimates

```{r estimate_table}
DT::renderDT({out$est})

```

