---
title: "Sport Harvest Estimator"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    source_code: embed
    theme:
      version: 4
      bootswatch: default
      #bg: "#D0D0D0"
      #fg: "#000000"
      navbar-bg: "#377F48"
      primary: "#377F48"
runtime: shiny
---

```{r setup_libraries, include=FALSE}
knitr::opts_chunk$set(comment = "#>", warning = F, echo = F)

library("flexdashboard")
library("tidyverse")
library("shiny")
library("patchwork")
library("gt")



```

```{r setup_db, include=FALSE}

# this is checking if the application is in a container
# (on the open internet) if it's not,
# it won't connect to the mongo db for storing user inputs
# pure web develop stuff has no influence on fish data


on_web_server = file.exists('/.dockerenv')

if (on_web_server){
  # load mongo, connect
  library("mongolite")
  user_ests <- mongo(collection='user_ests')
  
  # if saved dataframe doesn't exist yet populate it
  if (nrow(user_ests$find()) == 0) {
    user_ests$insert(
      expand_grid(
        area_code = c("05", "06", "07", "09", "10", "11", "12", "13", "81", "82"),
        timestep = as.character(1:5),
        est = 0, 
        wday_df = 0, wend_df = 0,
        wday_hpd = 0, wend_hpd = 0,
        rationale = ""
      )
      
    )
  }
}

```


```{r setup_theme, include=FALSE}
theme_set(theme_light())

pal_wkend <- c(wday = grey(0.3), wend = grey(0.8))

pal_m <- set_names(
  c("#241E33", "#18305A", "#004A70", "#006072", "#007373",
    "#158674", "#4B9976", "#74A97D", "#9CB98A", "#BECA92",
    "#E9DB87", "#FFE56C"), month.abb)

pal_ts <- set_names(
   c("#D2A554", "#626B5D", "#8C8F9E", "#858753", "#A4BADF"),
   paste(1:5, c("Jan-Jun","Jul","Aug","Sep","Oct-Dec"), sep = ":")
)

# this is on someone's personal computer 
# and not on a web server this table needs
# to be stored locally
if (on_web_server == FALSE){
  out <- reactiveValues(
    est = expand_grid(
      area_code = c("05", "06", "07", "09", "10", "11", "12", "13", "81", "82"),
      timestep = as.character(1:5),
      est = 0,
      wday_df = 0, wend_df = 0,
      wday_hpd = 0, wend_hpd = 0,
      rationale = ""
    )
  )
}



```


```{r pull_data}
# if on a web server pull data from a saved environment
# if it's not, pull from the sqlite datbase
if (on_web_server){
  load(file = "sport_harvest_estimator.RData")
} else {
  # connect to database
  con <- DBI::dbConnect(RSQLite::SQLite(), here::here('data/coho_harvest_estimator.db'))
  # pull necessary tables and view
  
  # puget sound creel information merged to regulation
  pssp <- DBI::dbReadTable(con, 'vw_estimates_by_regulation')
  
  
  # puget sound creel information merged to regulation
  regs <- DBI::dbReadTable(con, 'coho_regulations')
  
  
  # compilation of FRAM parameters from post-season and pre-season databases
  fs_ps_spt <- DBI::dbReadTable(con, 'coho_fram')
  
  # disconnect from database
  DBI::dbDisconnect(con)

}



# regulations need to be split out with weekday and weekday days counted
# for default inputs for days in upcomming fisheries fisheries

daily_regs <- regs %>%
  rowwise() %>%
  mutate(
    date = list(seq(as.Date(start_datetime), as.Date(end_datetime), by = '1 day'))
  ) %>%
  unnest_longer(date) %>%
  select(-start_datetime, -end_datetime) %>% # now a dataframe of daily dates by reg/area
  mutate(
    day_type = if_else(lubridate::wday(date) %in% c(1,6,7), 'WE', 'WD'), # weekday / weekend
    time_step = case_when(
      lubridate::month(date) %in% c(1,2,3,4,5,6) ~ 1L,
      lubridate::month(date)  == 7 ~ 2L,
      lubridate::month(date)  == 8 ~ 3L,
      lubridate::month(date)  == 9 ~ 4L,
      lubridate::month(date) %in% c(10, 11, 12) ~ 5L
      ),
    year = lubridate::year(date)
  )

# ugh, realized dates!!!!!!!!!!!!!!!!!!!!!!!!!!!! need planned/pamphlet!!!
regs_day_type_count <- daily_regs %>%
  count(catch_area_code, regulation_type_code, time_step, day_type, year) %>%
  pivot_wider(names_from = 'day_type', values_from = n)


# this is selecting the most recent number of WD/WE days 
# by area/timestep - defaults of the user interface

recent_regs <- regs_day_type_count %>%
  group_by(catch_area_code, time_step) %>%
  slice_max(order_by = year, n = 1) %>%
  ungroup()



```

About
============================================

This application supports the development of preseason coho FRAM marine sport fishery inputs.

The estimated harvest input is defined as the product of days fished $df$, anglers per day $apd$, and harvest-per-angler-per-day $hapd$ or: $harvest_{est} = df * apd * hapd$

Puget Sound dockside creel observations are used to inform the values for anglers per day $apd$, and harvest-per-angler-per-day $hapd$.

Annual sample distribution data are presented when the user selects a focal WDFW Puget Sound catch area and coho FRAM time step from the upper left radio buttons.

The anticipated season length determines the upper $df$ days fished slider, optionally differentiating between weekday and weekend "day types".

The mean, median (q50), and upper range (q95) of the anglers-per-day are used to inform the middle $apd$ slider. The raw values and a sample-rate based expansion are shown, with the slider choices based on latter.

The mean, median and interquartile range (IQR) are used to shape the lower harvest-per-angler-per-day $hapd$ slider. Adjusting the slider modifies the vertical reference line shown in the "PSSP kept obs" plots.

The harvest estimate (i.e., potential FRAM input) resulting from the current slider values is shown in the upper middle green box. 

After reaching an acceptable input, the user enters a brief description of the rationale and clicks the "Keep estimates" button, thereby preserving this text, the input, and the associated underlying $df$, $apd$, and $hapd$. After all desired inputs have been generated, a csv recording each area-timestep can be downloaded.

Coho kept
============================================

Area-Timestep-Estimates {data-height=250}
--------------------------------------------

### Area & Timestep

```{r buttons}
radioButtons("focal_area", label = "Focal Area",
             choices = c("05", "06", "07", "09", "10", "11", "12", "13", "81", "82"), selected = "05", inline = T, width = '100%')

radioButtons("focal_ts", label = "Focal Timestep",
             choiceValues = 1:5, 
             choiceNames = paste(1:5, c("Jan-Jun","Jul","Aug","Sep","Oct-Dec"), sep = ":"), 
             selected = 4, inline = T, width = '100%')

checkboxGroupInput("msf", label = "Regulation",
             choices = c('NS', 'MSF'), 
             selected = c('NS','MSF'), inline = T, width = '100%')



```

### Harvest estimates

```{r valuebox}
renderValueBox({
  valueBox(
    value = paste0("Weekday: ", round(wday_harvest()), " harvest"),
    icon = "fa-fish")
  })

renderValueBox({
  valueBox(
    value = paste0("Weekend: ", round(wend_harvest()), " harvest"),
    icon = "fa-fish")
  })

renderValueBox({
  valueBox(
    value = paste0("Total: ",
                   round(wday_harvest()+wend_harvest()), " harvest"),
    icon = "fa-fish")
  })

```

### Output

```{r keep_dl_buttons}


textInput("rationale", label = "Rationale")
actionButton("keep", label = "Keep estimates")


```

```{r keeper, echo=F}
observeEvent(
  input$keep,{
  est_row <- data.frame(
     area_code = input$focal_area,
     timestep = input$focal_ts,
     est = wday_harvest()+wend_harvest(),
     wday_df = input$wday_df, wend_df = input$wend_df,
     wday_hpd = input$wday_hpd, wend_hpd = input$wend_hpd,
     rationale = input$rationale
     )
  
  if (on_web_server){
    # pull df out of mongo
    est <- user_ests$find()
    
    # insert modified df
    est_insert <- rows_update(est, est_row, by = c("area_code", "timestep"))
    
    #remove previous data
    user_ests$remove('{}')
    
    #put new data in 
    user_ests$insert(est_insert)
    
  } else {
    # if not a websever use local
    out$est <- rows_update(out$est, est_row, by = c("area_code", "timestep"))
  }

   
   
})


```

```{r estimates prep}
# this chunk probably needs to be move to 
# an estimate prep file

# if taking the mean, 0's are important
# crcs use the dumb 'if there's a 0 i'll not add row' 
# strategy to their data... like it's the 70's and 
# we're still entering data on punch cards
# below is adding 0's back in to the means,
# being mindful that if dataset is incomplete
# to preserve NA's


# this is specifically dealing with cases where there was no coho catch in 
# entire fisheries  - nearly all time step 5/1
fisheries_no_catch <- pssp %>%
    filter(is.na(day_type)) %>%
    select(area_code = catch_area_code, regulation_type_code, start_datetime, end_datetime, starts_with('coho')) %>%
    mutate(
      across(starts_with('coho'), ~replace_na(.x, 0))
    ) %>%
    rowwise() %>%
    mutate(
      date = list(seq(as.Date(start_datetime), as.Date(end_datetime), by = '1 day'))
    ) %>%
    unnest_longer(date) %>% # explode stratum
    mutate(
    day_type = if_else(lubridate::wday(date) %in% c(1,6,7), 'WE', 'WD'), # weekday / weekend
    time_step = case_when(
      lubridate::month(date) %in% c(1,2,3,4,5,6) ~ 1L,
      lubridate::month(date)  == 7 ~ 2L,
      lubridate::month(date)  == 8 ~ 3L,
      lubridate::month(date)  == 9 ~ 4L,
      lubridate::month(date) %in% c(10, 11, 12) ~ 5L
      ),
    year = lubridate::year(date),
    source = 'CRC'
  )

# crc estimate cycles should be continuous up to 3/31 of the preceding management year
# we shouldn't have regulations into the future (this is what this tool does)
# to figure out the management year take the max year in the regulations
# say 2022 - everything after 2022-03-31 should either be intensive monitoring estimates
# or covered by crcs we don't have yet (NA)

# figure out most recent management year
# DO NOT PUT FUTURE DATES IN THE REGULATIONS FILE
management_year <- max(regs_day_type_count$year)

no_catch_no_estimate <- fisheries_no_catch %>%
  mutate(
    across(starts_with('coho'),
           ~if_else(date >= as.Date(glue::glue('{management_year}-04-01')), NA_real_, .x)
           )
  )

# this should be a list of fisheries where we don't have CRC estimates yet
# no_catch_no_estimate %>%
#   filter(is.na(coho_104_ad_ret)) %>%
#   count(area_code, start_datetime, end_datetime)

# this is a list of fisheries where the catch has been zero'd out (from CRC data)
# no_catch_no_estimate %>%
#   filter(!is.na(coho_104_ad_ret)) %>%
#   count(area_code, start_datetime, end_datetime) %>%
#   print(n=100)

# remove fisheries where there were no catch from estimates (added back in later dailyfied)
pssp <- pssp %>%
  filter(!is.na(day_type)) %>%
  mutate(date = as.Date(date)) %>%
  bind_rows(no_catch_no_estimate) %>%
  filter(year >= 2004) # crc weirdness with 2003

# now means make sense

# pre-process chart

# fram_chart <- pssp |>
#   group_by(year, time_step, area_code, source) |>
#   summarize(across(starts_with('coho'), sum, na.rm=T), .groups = 'drop') |>
#     rowwise() |>
#     mutate(
#       # apportion unk mark-status to ad/um for retention
#       coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / sum(coho_104_ad_ret,coho_104_um_ret , na.rm=T))),
#       coho_ad_k = replace_na(coho_104_ad_ret, 0),
#       coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret /    sum(coho_104_ad_ret,coho_104_um_ret, na.rm=T))),
#       coho_um_k = replace_na(coho_104_um_ret, 0),
#       total_retained_coho = sum(coho_104_ad_ret, coho_104_um_ret),
#       yr_type = paste0(year, '_pst_ests'),
#       yr = as.character(year),
#       ts = time_step,
#       type = 'ests'
#       ) |>
#     select(yr_type, yr, ts, type, area_code, var = source, val = total_retained_coho)     |>
#     bind_rows(fs_ps_spt)

```



```{r filtered_objects}
# these block of code will make dataframes reactive objects.
# i.e. when the variable pssp_at() is called farther down
# this code it's referencing the reactive object below,
# as the user changes the inputs of area, timestep, and years
# the pssp_at() object executes the code below, dynamically filtering the
# pssp dataframe to the users input represented by pssp_at() 

pssp_at <- reactive({
  filter(pssp, 
         area_code == input$focal_area,
         time_step == input$focal_ts, 
         year %in% input$years)
})

fram_at <- reactive({
    filter(fs_ps_spt,
         area_code == input$focal_area, 
         ts == input$focal_ts, 
         yr %in% input$years
    )
}) 

recent_regs_at <- reactive({
  filter(recent_regs,
         catch_area_code == input$focal_area, 
         time_step == input$focal_ts)
})


# consider putting these filters in the reactive containers 
pssp_at_reg <- reactive({
  filter(pssp_at(), regulation_type_code %in% input$msf)
})

fram_at_reg <- reactive({
  fram_at() %>%
    mutate(
      selector = if_else(var == 'MSFQuota', 'MSF', 'NS') 
    ) %>%
    filter(
         selector %in% input$msf
         ) %>% select(-selector)
}) 

# pssp %>%
#   group_by(day_type) %>%
#   summarize(across(starts_with('coho'), mean, na.rm=T))

```



```{r default_controls}
# these are the controls for the defaults for the weekday and weekend
# inputs. the days are pulled from the most recent fishery by area/timestep
# harvest is the mean selected from the year checkboxes

# application supplied weekend/weekdays
observeEvent(ignoreInit = TRUE,
             c(input$focal_area, input$focal_ts), {


  wd_days <- recent_regs_at() %>%
  # calculate most recent regulation weedays
    .$WD

  updateNumericInput(inputId = 'wday_df', value = wd_days )


  we_days <- recent_regs_at() %>%
    .$WE

  updateNumericInput(inputId = 'wend_df', value = we_days )

})

# application supplied harvest per day
observeEvent(ignoreInit = TRUE,
             c(input$focal_area,
               input$focal_ts,
               input$years,
               input$msf), {

  # catch a corner case, if user input doesn't have fisheries to
  # take a mean of, turn off the regulation filtering
  if(nrow(pssp_at_reg()) == 0){
    showModal(modalDialog(
      title = 'No Usuable Data',
      'There are no previous data to calculate mean harvest by day. 
      Defaulting to whatever data available')
    )
    
    updateCheckboxGroupInput(inputId = 'msf', selected = c('NS', 'MSF'))
    updateCheckboxGroupInput(inputId = 'years', selected = 2004:2022)
  } else {
    
    mean_harvest <- pssp_at_reg() |>
      mutate(
        # take care of NA's
        across(starts_with('coho'), ~replace_na(.x, 0)),
        # apportion unk mark-status to ad/um for retention
        coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / (coho_104_ad_ret + coho_104_um_ret ))),
        coho_ad_k = replace_na(coho_104_ad_ret, 0),
        coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret /  (coho_104_ad_ret + coho_104_um_ret ))),
        coho_um_k = replace_na(coho_104_um_ret, 0),
        total_retained_coho = coho_ad_k + coho_um_k
      ) |>
      group_by(time_step, area_code, day_type) %>%
      summarize(mean_daily_harvest = mean(total_retained_coho, na.rm=T), .groups = 'drop') %>%
      pivot_wider(names_from = day_type, values_from  = mean_daily_harvest) %>%
      ungroup()




  wd_mean_harvest <- mean_harvest %>%
    .$WD

  we_mean_harvest <- mean_harvest %>%
    .$WE

  updateNumericInput(inputId = 'wend_hpd', value = round(we_mean_harvest))

  updateNumericInput(inputId = 'wday_hpd', value = round(wd_mean_harvest))


  }

})



```


Controls {data-height=200}
--------------------------------------------

### Weekday assumptions
```{r wkday_sliders}



numericInput("wday_df", label = "Weekday Days fished",
            min = 0, max = 365, value = 20, step = 1, width = '100%')

# numericInput("wday_apd", label = "Weekday Anglers per day",
#             min = 0, max = 2000, value = 0, step = 5, width = '100%')

numericInput("wday_hpd", label = "Mean Weekday Day Harvest",
            min = 0, max = 4, value = 0, step = 0.01, width = '100%')

wday_harvest <- reactive({
  input$wday_df * input$wday_hpd
})

```

### Weekend assumptions

```{r wkend_sliders}
numericInput("wend_df", label = "Weekend Days fished",
            min = 0, max = 365, value = 20, step = 1, width = '100%')

# numericInput("wend_apd", label = "Weekend Anglers per day",
#             min = 0, max = 2000, value = 0, step = 5,  width = '100%')

numericInput("wend_hpd", label = "Mean Weekend Day Harvest",
            min = 0, max = 4, value = 0, step = 0.01,  width = '100%')

wend_harvest <- reactive({
  input$wend_df * input$wend_hpd
})


```





Years {data-height=125}
--------------------------------------------

### Year selection

```{r years_checkboxes}
checkboxGroupInput("years", "Focal years", inline = T, width = '100%',
                   choices = sort(unique(pssp$year)), 
                   selected = sort(unique(pssp$year))
                   )
```


Data {.tabset data-height=800}
--------------------------------------------

### FRAM
```{r gg_fram}
renderPlot({
  fram_ests <- pssp_at_reg() |>
    group_by(year, time_step, area_code, source) |>
    summarize(across(starts_with('coho'), sum, na.rm=T), .groups = 'drop') |>
    rowwise() |>
    mutate(
      # apportion unk mark-status to ad/um for retention
      coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / sum(coho_104_ad_ret,coho_104_um_ret , na.rm=T))),
      coho_ad_k = replace_na(coho_104_ad_ret, 0),
      coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret / sum(coho_104_ad_ret,coho_104_um_ret, na.rm=T))),
      coho_um_k = replace_na(coho_104_um_ret, 0),
      total_retained_coho = sum(coho_104_ad_ret, coho_104_um_ret),
      yr_type = paste0(year, '_pst_ests'),
      yr = as.character(year),
      ts = time_step,
      type = 'ests'
      ) |>
    select(yr_type, yr, ts, type, area_code, var = source, val = total_retained_coho)     |>
    bind_rows(fram_at_reg())
  


  fram_ests |> ggplot(aes(yr_type, val, fill = var)) +
    geom_col(position = position_stack(), alpha = 0.7) +
    #geom_col(data = filter(fram_ests, type == "ests"), position = position_stack(), alpha = .6) +
    #geom_col(data = filter(fram_ests, type == "pst"), position = position_stack(), alpha = 1) +
    geom_hline(yintercept = wday_harvest() + wend_harvest()) +
    scale_y_continuous("FRAM inputs", sec.axis = dup_axis(), labels = scales::comma) +
    scale_fill_manual(values = c(
      "Quota" = "#D2A554",
      "MSFQuota" = "#A4BADF",
      "CRC" = "#04a43c",
      "CREEL" = "#014218"
      )) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ","))
    ) +
    theme(
      text = element_text(size = 20),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
      )
})

```

### Daily Harvest
```{r gg_daily}
renderPlot({
  day_plot <- pssp_at_reg() |>
      mutate(
        # take care of NA's
        across(starts_with('coho'), ~replace_na(.x, 0)),
        # apportion unk mark-status to ad/um for retention
        coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / (coho_104_ad_ret + coho_104_um_ret ))),
        coho_ad_k = replace_na(coho_104_ad_ret, 0),
        coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret /  (coho_104_ad_ret + coho_104_um_ret ))),
        coho_um_k = replace_na(coho_104_um_ret, 0),
        total_retained_coho = coho_ad_k + coho_um_k
      )
  
  day_plot |> 
    ggplot(aes(factor(year), total_retained_coho, fill = day_type, fill = day_type )) +
    geom_boxplot() +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ",")),
      y = 'Total Retained Coho',
      x = 'Day Type'
    ) + 
    theme(
      text = element_text(size = 20)
      )
    
})
```

### FRAM Performance

```{r fram_performance}

renderPlot({
  ests <- pssp_at() %>%
    mutate(
      # take care of NA's
      across(starts_with('coho'), ~replace_na(.x, 0)),
      # apportion unk mark-status to ad/um for retention
      coho_ad_k = coho_104_ad_ret + (coho_104_uk_ret  * (coho_104_ad_ret / (coho_104_ad_ret + coho_104_um_ret ))),
      coho_ad_k = replace_na(coho_104_ad_ret, 0),
      coho_um_k = coho_104_um_ret + (coho_104_uk_ret * (coho_104_um_ret /  (coho_104_ad_ret + coho_104_um_ret ))),
      coho_um_k = replace_na(coho_104_um_ret, 0),
      total_retained_coho = coho_ad_k + coho_um_k
    ) |>
    group_by(time_step, area_code, year) %>%
    summarize(total_retained_coho = sum(total_retained_coho, na.rm=T), .groups = 'drop')
  
  fram <- fram_at() %>%
    mutate(
      year = as.integer(yr),
      time_step = ts
    ) %>%
    group_by(year, time_step, area_code) %>%
    summarize(predicted_retained_coho = sum(val, na.rm=T), .groups = 'drop')
  
  fram %>%
    inner_join(ests) %>%
  mutate(
    percent = (total_retained_coho - predicted_retained_coho) / predicted_retained_coho,
    type = percent < 0
  ) %>%
    ggplot(aes(factor(year), percent, fill = type)) +
    geom_col(alpha = 0.8) +
    labs(
      subtitle = paste("Area", input$focal_area, "during", paste(input$focal_ts, collapse = ",")),
      x = NULL,
      y = 'Percent Difference of Modeled Quota'
    ) +
    scale_y_continuous(labels = scales::percent) +
    theme(
      axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1),
      text = element_text(size = 20),
      legend.position = 'none'
    )
  
  
})
```



### Saved Inputs
```{r, download_link}
downloadLink("download", label = "Download csv")
if (on_web_server){
  output$download <- downloadHandler(
    filename = function() {
      paste("coho_sport_harvest_est_", Sys.Date(), ".csv", sep="")
    },
    content = function(file) {
      write_csv(user_ests$find(), file)
    }
  )
} else {
  output$download <- downloadHandler(
    filename = function() {
      paste("coho_sport_harvest_est_", Sys.Date(), ".csv", sep="")
    },
    content = function(file) {
      write_csv(out$est, file)
    }
  )
}

```

```{r ests_table}
if (on_web_server) {
  # poll database
  users_inputs <- reactivePoll(1000, session,
                               
                               # check if something changed
                               checkFunc = function() {
                                 
                                 ests <- user_ests$find()
                                 
                                 
                                 sum(ests$est, ests$wday_df, 
                                     ests$wend_df, ests$wday_hpd, 
                                     ests$wend_hpd)
                               },
                               valueFunc = function() {
                                 # pull new data
                                 user_ests$find()
                               }
                               
                               
  )
  # display table from database
  renderTable({
    users_inputs()
  })
} else{
  # otherwise display local
  renderTable({
    out$est
  })
}


```

```{css est_table}
#section-saved-inputs
  .chart-shim {
    overflow-y: scroll;
    }
```



